---
title: "PageRank"
author: "Karan Pillay"
date: "November 24th, 2019"
output:
 html_document:
    toc: true
---

a). The corresponding transition matrix A is:

For Diagram a)
```{r setup, transitionmatrixa}
#format of matrix function is (col1^T,col2^T, etc)
diagram_a=matrix(c(0,1/3,1/3,1/3, 0, 0, 1/2, 1/2, 1, 0, 0, 0, 1/2, 0, 1/2, 0), nrow=4)
diagram_a
```
The above matrix is the transition matrix for diagram a.


For Diagram b)
```{r setup, transitionmatrixab}
#format of matrix function is (col1^T,col2^T, etc)
diagram_b=matrix(c(0,1/3,1/3,1/3, 0, 0, 0, 0, 1/2, 1/2, 1, 0, 0, 0, 0, 0, 1/2, 1/2, 0, 0, 1/5, 1/5, 1/5, 1/5, 1/5), nrow=5)
diagram_b
```
The above matrix is the transition matrix for diagram b.



The google matrix G is:

For diagram a.
```{r setup, google_a}
#Note: The following does not need to be touched or adjusted as long as the matrices above are manually entered. 

p=.15
one_minus_p=1-p
num_rows_a=nrow(diagram_a)
one_over_k_a=1/num_rows_a
matrix_of_ones_a<-matrix(1,nrow=num_rows_a,ncol=num_rows_a)
G_a=(one_minus_p*diagram_a)+(p*one_over_k_a*matrix_of_ones_a)
G_a
```
The above matrix is the Google Matrix of diagram a.



For diagram b.
```{r setup, google_b}
#Note: The following does not need to be touched or adjusted as long as the matrices above are manually entered. 

p=.15
one_minus_p=1-p
num_rows_b=nrow(diagram_b)
one_over_k_b=1/num_rows_b
matrix_of_ones_b<-matrix(1,nrow=num_rows_b,ncol=num_rows_b)
G_b=(one_minus_p*diagram_b)+(p*one_over_k_b*matrix_of_ones_b)
G_b
```
The above matrix is the Google Matrix of diagram b.



c).Does either graph have a dangling node? If so, which one(s)?

Yes, diagram b has a dangling node. Node 5 on diagram b is a dangling node. 

d). Does either graph have a disconnected node? If so, which one(s)?

There does not exist a disconnected node in either of the two diagrams. 

e). The following represent the PageRank of each page in the set respectively. 

PageRank for diagram a:
```{r setup, pageRank_a}
v_a=eigen(G_a)
v_a$vectors
v_a$vectors[,1]/sum(v_a$vectors[,1])

```
The above vector is the PageRank for diagram a.


PageRank for diagram b:
```{r setup, pageRank_b}
v_b=eigen(G_b)
v_b$vectors
v_b$vectors[,1]/sum(v_b$vectors[,1])
```
The above vector is the PageRank for diagram b.