---
title: "Singular Vector Decomposition"
author: "Karan Ravindran Pillay"
date: "11/20/2019---12/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Statement: Given a set of manually classified digits (the training set), classify a set of unknown digits (the test set) using SVD method.
## Data Set
#### Loading the Data...
```{r}
trI <- read.csv("trainInput.csv",header = FALSE)
trO <- read.csv("trainOutput.csv", header = FALSE)
teI <- read.csv("testInput.csv",header = FALSE)
teO <- read.csv("testOutput.csv", header = FALSE)
```

#### Reorganizing the Data...
```{r}
trainInput <- t(trI)
trainOutput <- t(trO)
trainData <- cbind(trainInput,trainOutput)
```
## Methodology
#### Form the Matrix *A[[i]]* for Digit i-1(Such That Each Row in *A[[i]]* Represents an Image of Digit i-1). <br>
#### Determine the Singular Value Decomposition of Each *A[[i]]*(and store it in *S[[i]]*). <br>
#### Retrieve the First 20 Singular Images of Each Digit(and store them in *V[[i]]*)
```{r}
A <- list()
S <- list()
V <- list()
for (i in 1:10){
  A[[i]] <- subset(trainData, trainData[,257] == i-1)[,-257]
  S[[i]] <- svd(A[[i]])
  V[[i]] <- S[[i]]$v[,1:20]
}
```

#### Express Test Images as a Linear Combination of the First k=20 Singular Images of Each Digit.<br>
$$testImage ≈ V[[i]]x$$ <br>

#### Compute the Distance Between Test Images and Their Least Square Approximations and Store in *n[i]*. <br>

#### Classify Each Test Image to be the *digit[j]* Corresponding to the Smallest Residual.

```{r}
library(MASS)
n <- c()
digit <- c()
for (j in 1: ncol(teI)){
  for (i in 1:10) {
  n[i] <- norm(V[[i]] %*% (ginv(V[[i]]) %*% teI[,j]) - teI[j], '2')
  }
  digit[j] <- which.min(n)-1
}
#digit
```

#### Calculate the Overall Correct Classification Rate.
```{r}
count <- 0
for(i in 1:ncol(teO)){
  if(digit[i] == teO[i]){
    count <- count + 1
  }
}
corrRate <- count/ncol(teO)
corrRate
```

#### Calculate the Correct Classification Rate for Each Digit in a Confusion Matrix.
```{r, warning = FALSE, message = FALSE}
library(knitr)
library(kableExtra) 
library(dplyr)
predict <- digit
reference <- t(teO)
confMat <- as.matrix(table(reference,predict))
confMat %>%
  kable(format = "html", caption = "Classification Correct Rate: Reference | Predict -->") %>%
  kable_styling(font = 20) %>%
  row_spec(0, bold = T, color = "tomato")
  
rate <- c()
for (i in 1:10){
  rate[i] <- confMat[i,i]/rowSums(confMat)[i]
}

df <- cbind(c(0:9),rate)
colnames(df) <- c("Digit", "CorrRate")
df %>%
  kable(format = "html", caption = "Classification Correct Rate") %>%
  kable_styling(font = 20) %>%
  row_spec(0, bold = T, color = "tomato")
```


### Summary
For R Command:<br>
a. n[3] != n[3] without firstly define n<br>
b. dataframe[,1] != dataframe[1]. as.matrix(df[,1]) == df[1]<br>
c. which.min() is for retrieving index of minimum.<br>
For SVD Application: <br>
a. Model Comes From V <br>
b. Using First a Few Concepts Not Features to Represent the Whole <br>
c. Least Square Method is for Comparing and Classifying <br>
For Accuracy Result Interpretation: <Br>
```{r}
cat("The Highest Predict Accuracy is for Digit: ", which.max(rate)-1,
    "\nThe Lowest Predict Accuray is for Digit: ", which.min(rate)-1)
```
Possiblely because: 
```{r}
for(i in 1:10){
  cat("Digit", i-1, "'s Matrix Dimension is", dim(A[[i]]),"\n")
}
```
From which, we can see:
```{r}
num_sample <- c()
for(i in 1:10){
  num_sample[i] <- nrow(A[[i]])
  #cat("Digit", i-1, "'s Has", num_sample[i]," Samples\n")
}
cat("The Digit that Has the Largest Sample Number in Training Dataset is ", which.max(num_sample)-1,
    "\nThe Digit that Has the Lowest Sample Number in Training Dataset is: ", which.min(num_sample)-1)
```

```{r}
df <- cbind(num_sample, rate)
```

```{r}
library(psych)
options(repr.plot.width=5, repr.plot.height=5)
pairs.panels(df[,c("num_sample", "rate")], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots 
             ellipses = TRUE # show correlation ellipses
)
```


#### Thus, we confirm the previous intuitive believe, that the more sample that the model build with, the more accuracy the model may work with.

